---
title: "QTM 6300 [Babson College]: Assignment 2"
author: "Eric W. Chan"
date: "2024"
format: html
theme: lumen
self-contained: true
editor: source
toc: true
toc-depth: 2
---

We will be using the BostonHousing.csv data. Please familiarize yourself with the data set first using the data dictionary prior to doing the below. Make sure to follow these directions:

* Please submit a single R script.
* Also, specific to this assignment, feel free to ask each other for help as I would like you to successfully write an R script, but please submit your own script.
you will receive credit for this assignment as long as it is good effort.
* Generally, it is good practice to comment using "#" exactly what you are doing, but especially what number you are answering. 
* This is due before next class. Late Assignments (without prior approval) will only be accepted at most 24 hours after the due date/time. Unexcused late assignments will incur an automatic 25% loss of points.

# Questions



1. At the top of your file, clear your global environment (the memory of your data frames and other objects) by using the function rm(list=ls()).

```{r}
rm(list=ls())
```

2. Set your working directory in the gray space provided. 

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/Users/diana/Babson MSBA/Machine learning")
```

3. Import the BostonHousing.csv data set. Name this data frame as "housing." Write the proper code.

```{r}
housing <- read.csv("BostonHousing.csv")
```

4. View the data using the View() and head() functions. Write the proper code.

```{r}
#View(housing)
head(housing) 
```

5. What is the data type of _AGE_? Use a function in R and answer the question  in space provided.

Answer: numeric

```{r}
str(housing$AGE)
```

6. What is the data type of _CHAS_? Use a function in R and answer the question in space provided.

Answer: integer 

```{r}
str(housing$CHAS)
```


7. Are there any variables with the data type "CHAR"? Use a function in R and answer the question.

Answer: No 

```{r}
is.character(housing)
```


8. Are there any variables that looks like it might be appropriate to be converted to a factor variable? Which ones and why?

Answer: Yes, 3 variables can be converted to a factor variable: CHAS, ISHIGHVAL and RAD is tricky because it is measuring from a scale of 1 (being the nearest) to 25 being the farest and only takes distance into consideration. For example, you could be near a highway, but the way to access it requires entering a private lot, so you have to go longer to enter.

```{r}

```

9. What is the data point located in row 233, column 6?

Answer: 8.337 average number of rooms per dwelling in house 233

```{r}
housing[233,6]
```


10. What is the data point located in row 378, column 4?

Answer: Is not near charles river, the data point is 0 

```{r}
housing [378,4]
```

11. Use the proper R function to view only rows 202 through 204.


Answer: 

```{r}
housing[202:204, ]
```

12. Keep only data for the columns _CRIM_, _CHAS_, _AGE_, and _MEDV_. Create a new data frame for this set of columns called "housing_subset."

Answer: 

```{r}
housing_subset <- housing[,c("CRIM","CHAS", "AGE" , "MEDV")]
#View(housing_subset)
```

13. Using housing_subset, Plot the relationship between the _CRIM_ and _MEDV_ variables.

Answer:

```{r}
library(ggplot2)
ggplot(data = housing_subset, aes(x=CRIM, y=MEDV))+geom_line()
```

14. Using housing_subset, typecast _CHAS_ as a factor variable.

Answer: 

```{r}
housing_subset$CHAS = as.factor(housing_subset$CHAS)
```

15. Using housing_subset, how many locations are are within census tract bounds of the river (see the _CHAS_ variable). Use the table() function to see this.

Answer: 35 locations are within census tract bounds of the river

```{r}
table(housing_subset$CHAS) 
```
